
\chapter*{Acknowledgments}
\markboth{Acknowledgments}{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}
\vspace{1.7cm}
%
My PhD was an incredible journey. Despite the deep passion that was driving my daily work, the last four years were a roller coaster of emotions. Moments of (often naive) excitement were quickly replaced by episodes of undue pessimism. Yet, the whole experience was deeply joyful, even in the tough moments, and this was thanks to the constant support that I received from the people around me.

First, I want to thank Nicola M., for his invaluable lessons and for always having faith in me (and in Koopmans!). When I was struggling, his unconditional trust allowed me to work serenely and without any pressure.

A big thanks to Nicola C., for being more than a supervisor. His meticulousness, passion for understanding and knowledge of the field (without forgetting his ``pills'' of wisdom) were always a model to me, and would constantly raise the bar pushing me to learn more and more.

Thanks also to Edward, for the organization and the high-level Python expertise that he brought to the team, and especially for the patience and the time he dedicated to me. I learned really a lot from him.

I thank my colleagues from THEOS, for the very nice environment that they contributed to create in the group. But a special thanks goes to Norma, Francesco and Tommaso. Sharing this experience together and supporting each other was one of the best parts. And I cannot avoid thanking Norma also for being the ``beta reader'' of this manuscript, and Tommaso for all the scientific battles that we had throughout these years. Most of the understanding and the knowledge that I tried to condense in this thesis are the result of those discussions.

Thanks to my friends ``from Castelli'' whom, despite the distance and the difficulties of the recent times, I knew I could always count on.

I thank a lot my family (including the Swiss-Dutch one), for always being there for me. If I am here today is in a large part thanks to them.

Finally, the biggest thanks goes to Amanda for listening, comforting and motivating me more than anybody else. You always had the right words (not an easy task with me!), and you were always making me happy, even when Koopmans were not.